Explicit B-spline Regularization in Diffeomorphic Image Registration

### 2013-09-07 paper.tex
ENH:  Adding files.

Parallel to this branch of algorithmic progress are the informally 
denoted ``dense transforms''
perhaps best exemplified by Thirion's seminal contribution \citep{thirion1998}.
Relationships with earlier elastic \citep{bajcsy1989,gee1993} and fluid \citep{christensen1996} registration methods are elucidated in 
the works of \cite{bro-nielsen1996,pennec1999} who observe that
Gaussian smoothing, characteristic of Demons, of the update or displacement
field is a greedy approximation for solving the partial differential equations governing
the physics of an elastic or fluid deformation.  However, the use of such 
approximations comes with the caveat that physical properties, such as topological
regularity,
are no longer guaranteed.

It is interesting to note that within this context, traditional FFD algorithms
can be viewed as a type of fluid-like Demons approach 
where, rather than produced by Gaussian convolution, smooth fields are produced using B-splines for roughly approximating the solution to the viscous fluid system.  This was hinted at in our earlier work \citep{tustison2009} where we showed that fitting the update field to a B-spline object using a fast approximation routine \citep{tustison2006} is equivalent to a preconditioning of the standard gradient used in gradient descent-based FFD 
optimization.  Further details can be gleaned from the ITK implementation of this work%
\footnote{
http://www.itk.org/Doxygen/html/classitk\_1\_1BSplineSmoothingOnUpdateDisplacementFieldTransform.html
}
which permits both B-spline smoothing on the update (``viscous'') and total (``elastic'') 
displacement fields at each iteration (cf
analogous Gaussian, i.e. Demons, implementation%
\footnote{
http://www.itk.org/Doxygen/html/classitk\_1\_1GaussianSmoothingOnUpdateDisplacementFieldTransform.html
}).

>>>>

Parallel to this branch of algorithmic progress are the informally 
denoted ``dense transforms''
perhaps best exemplified by Thirion's seminal contribution \citep{thirion1998}.
Relationships with earlier elastic \citep{bajcsy1989,gee1993} and fluid \citep{christensen1996} registration methods are detailed in 
the works of \cite{bro-nielsen1996} and \cite{pennec1999} who observe that
Gaussian smoothing, characteristic of Demons, of the update or displacement
field is a greedy approximation for solving the partial differential equations governing
the physics of an elastic or fluid deformation.  However, the use of such 
approximations comes with the caveat that physical properties, such as topological
regularity, are no longer guaranteed.

It is interesting to note that within this context, traditional FFD algorithms
can be viewed as a type of fluid-like Demons approach 
where, rather than produced by Gaussian convolution, smooth fields are produced using B-splines for roughly approximating the solution to the viscous fluid system.  This was hinted at in our earlier work \citep{tustison2009} where we showed that fitting the update field to a B-spline object using a fast approximation routine \citep{tustison2006} is equivalent to a preconditioning of the standard gradient used in gradient descent-based FFD optimization.  This preconditioning used to reduce the

Further details can be gleaned from the ITK implementation of this work%
\footnote{
http://www.itk.org/Doxygen/html/classitk\_1\_1BSplineSmoothingOnUpdateDisplacementFieldTransform.html
}
which permits both B-spline smoothing on the update (``viscous'') and total (``elastic'') 
displacement fields at each iteration (cf
analogous Gaussian, i.e. Demons, implementation%
\footnote{
http://www.itk.org/Doxygen/html/classitk\_1\_1GaussianSmoothingOnUpdateDisplacementFieldTransform.html
}).
